Reasoning About a Machine with Local Capabilities

### 2019-02-25 toplas/paper.tex
[TOPLAS] Update syntax figure to be less dense
[TOPLAS] Rephrase local capability explanation

\begin{figure}[b]
\[\arraycolsep=1.4pt
  \begin{array}{rlrcl l rlrcl}
    \addr   &\in& \Addrs &\isdef& \nats &\phantom{mak}                                & r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
    w &\in&\Words &\isdef& \ints + \Caps &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words \\
    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
    & & & &\exec\mid \entry\mid \rwx\mid \rwlx &                                             & \Phi &\in & \ExecConfs   &\isdef& \Regs \times \Heaps \\
    \gl&\in&\Globals & ::= &\glob \mid \local &                                              & \ms  &\in & \MemSegments &::= & \Addrs \parfun \Words \\
    &&\Confs &::=& \multicolumn{6}{l}{\ExecConfs + \{\failed \} + \{\halted\} \times \Heaps}\\
     && \Caps  &\isdef & \multicolumn{6}{l}{\left\{
                         ((\perm,\gl),\start,\addrend,\addr)\mid
                         \start,\addr\in\Addrs,\addrend \in
                         \Addrs\cup \{\infty\}\right\}}
  \end{array}
\]
  \begin{equation*}
  \begin{array}{rcl}
    \rv    &\in& \ints + \RegName \\
    i      &::=& 
                 \jmp{\lv} \mid 
                 \jnz{\lv}{\rv} \mid
                 \move{\lv}{\rv} \mid 
                 \load{\lv}{\hv} \mid 
                 \store{\hv}{\rv} \mid
                 \lt{\lv}{\rv}{\rv} \mid
                 \plus{\lv}{\rv}{\rv} \mid 
                 \minus{\lv}{\rv}{\rv} \mid \\
          &   &  \lea{\lv}{\rv} \mid 
                 \restricttwo{\lv}{\rv} \mid
                 \subseg{\lv}{\rv}{\rv} \mid  
                 \isptr{\lv}{\rv} \mid 
                 \getl{\lv}{\lv} \mid
                 \getp{\lv}{\lv} \mid
                 \getb{\lv}{\lv} \mid\\
          &   &  \gete{\lv}{\lv} \mid 
                 \geta{\lv}{\lv} \mid 
                 \fail \mid
                 \halt 
  \end{array}
\end{equation*}
  \caption{The syntax of our capability machine assembly language.}
  \label{fig:syntax}
\end{figure}

@@

Finally, the $\store{}{}$ instruction updates the memory to the requested value
if the capability has write authority for the requested location. However, the
instruction is complicated by the presence of \emph{local capabilities} modeled
after the ones in the CHERI processor~\citep{Watson2015Cheri}. Basically, local
capabilities are special in that they can only be kept in registers, i.e.\ they
cannot be stored to memory. This means that local capabilities can be
\emph{temporarily} given to an adversary, for the duration of an invocation: if
we make sure to clear the capability from the register file after control is
passed back to us, they will not have been able to store the capability.
However, there is one exception to the rule above: local capabilities can be
stored to memory for which we have a capability with write-local authority
(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
where register contents can be stored, including local capabilities. As long as
all capabilities with write-local authority are themselves local and the stack
is cleared after control is passed back by the adversary, we will see that this
does not break the intended behavior of local capabilities.

@@

Plenty of other papers\lau{If there are plenty, then I guess we should cite more than one?}
enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
This line of work uses a quite different attacker model than us.
They assume an attacker that is not able to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
By checking the address of every indirect jump and using memory access control to prevent overwriting code, this work enforces what they call control-flow integrity formalized as the property that every jump will follow a legal path in the control-flow graph.
As discussed in \sectionname~\ref{sec:discussion}, such a property ignores temporal properties and seems hard to use for reasoning.

>>>>

\begin{figure}[b]
\[\arraycolsep=1.4pt
  \begin{array}{rlrcl}
    \addr   &\in& \Addrs &\isdef& \nats \\
    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid\exec\mid \entry\mid \rwx\mid \rwlx  \\
    \gl&\in&\Globals & ::= &\glob \mid \local \\
    c &\in& \Caps  &\isdef & \left\{
                         ((\perm,\gl),\start,\addrend,\addr)\mid
                         \start,\addr\in\Addrs,\addrend \in
                         \Addrs\cup \{\infty\}\right\}\\
    w &\in&\Words &\isdef& \ints + \Caps \\
    r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
    \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words \\
    m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
    \Phi &\in & \ExecConfs   &\isdef& \Regs \times \Heaps \\
    \ms  &\in & \MemSegments &\isdef& \Addrs \parfun \Words \\
    &&\Confs &\isdef& \ExecConfs + \{\failed \} + \{\halted\} \times \Heaps
  \end{array}
\]
  \begin{equation*}
  \begin{array}{rcl}
    \rv    &\in& \ints + \RegName \\
    i      &::=& 
                 \jmp{\lv} \mid 
                 \jnz{\lv}{\rv} \mid
                 \move{\lv}{\rv} \mid 
                 \load{\lv}{\hv} \mid 
                 \store{\hv}{\rv} \mid
                 \lt{\lv}{\rv}{\rv} \mid
                 \plus{\lv}{\rv}{\rv} \mid 
                 \minus{\lv}{\rv}{\rv} \mid \\
          &   &  \lea{\lv}{\rv} \mid 
                 \restricttwo{\lv}{\rv} \mid
                 \subseg{\lv}{\rv}{\rv} \mid  
                 \isptr{\lv}{\rv} \mid 
                 \getl{\lv}{\lv} \mid
                 \getp{\lv}{\lv} \mid
                 \getb{\lv}{\lv} \mid\\
          &   &  \gete{\lv}{\lv} \mid 
                 \geta{\lv}{\lv} \mid 
                 \fail \mid
                 \halt 
  \end{array}
\end{equation*}
  \caption{The syntax of our capability machine assembly language.}
  \label{fig:syntax}
\end{figure}

@@

Finally, the $\store{}{}$ instruction updates the memory to the argument value
if the capability has write authority for the specified location. However, the
instruction is complicated by the presence of \emph{local capabilities} modeled
after the ones in the CHERI processor~\citep{Watson2015Cheri}. At a high-level, local
capabilities are special in that they can only be kept in registers, i.e.\, they
cannot be stored to memory. This means that local capabilities can be
\emph{temporarily} given to an adversary, for the duration of an invocation. If
we make sure to clear the capability from the register file after control is
passed back to us, the adversary is unable to store the capability.
However, there is one exception to the rule above: local capabilities can be
stored to memory for which we have a capability with write-local authority
(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack
where register contents can be stored, including local capabilities. As long as
all capabilities with write-local authority are themselves local and the stack
is cleared after control is passed back by the adversary, we will see that this
does not break the intended behavior of local capabilities.

@@

Plenty of other papers\lau{If there are plenty, then I guess we should cite more than one?}
enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
This line of work uses a quite different attacker model than us.
They assume an attacker that is unable to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
By checking the address of every indirect jump and using memory access control to prevent overwriting code, this work enforces what they call control-flow integrity formalized as the property that every jump will follow a legal path in the control-flow graph.
As discussed in \sectionname~\ref{sec:discussion}, such a property ignores temporal properties and seems hard to use for reasoning.
